Manuscript draft

### 2019-09-09 manuscript/manuscript draft.Rmd
update manuscript draft

| Variable | Definition | Components included | Methodologies used | Number of records | Number of plots |
|--------------|-------------------|-------------|------------------------|-------|-------|
| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 257 | 80 |
| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 93 | 78 |
| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 258 | 166 |
| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 94 |
| ANPP~woodystem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 266 | 252 |
| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 107 | 87 |
| BNPP~fineroot~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 90 | 77 |
| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 65 | 43 |

@@

*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes. To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component. The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux. Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015]. Stacked plots were generated for:  
1. GPP = NPP + R~auto~  
2. NPP = ANPP + BNPP  
3. ANPP = ANPP~foliage~ + ANPP~woody stem~  
4. Total belowground carbon flux = BNPP + R~root~

*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody stem~. For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated. The ratios were regressed against latitude and climate variables, using the linear model specified above. Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions. To account for this, models were re-run using only data from sites $\leq$ 1000m.

@@

[*Relationships and differences among fluxes- *] In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig. 2). That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.

@@

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine root~, ANPP, ANPP~woody stem~, and R~root~ (fig.4). There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.

@@

We found a significant relationship between productivity and potential evapotranspiration for all fluxes. ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig. 5). We find strong evidence for a saturation point or peak with PET: productivity tends to increase at values below 1000mm, before saturating between 1200 and 1700mm. There is evidence that productivity begins to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of productivity for all fluxes. BNPP~fine root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig. 5). Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa. At values above 0.8 kPa, productivity began to decrease.

@@

```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors: mean annual temperature; temperature seasonality; annual temperature range; and vapour pressure. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("combined_plots.png")
```

*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity. We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~. Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases (fig. 5). ANPP~foliage~, ANPP~woody stem~ and R~auto~ decrease linearly with temperature seasonality. Temperature seasonality is strongly correlated with annual temperature range, and, as expected, all fluxes show almost identical responses to it. Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

@@

*Within the growing season, how does productivity vary with climate?* Within growing season months, we find that climate has a much weaker effect on productivity. For each of temperature, precipitation, PET, and solar radiation, we find a small effect of climate for certain carbon fluxes. There is a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP_foliage (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05). Productivity increases with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine root~ (R^2^ = 0.11, p<0.05), and ANPP~woody stem~ (R^2^ = 0.06, p<0.05).

@@

| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | MAT*MAP |
|----------|----------|----------|----------|--------------|----------|
| R~auto~ | 0.65 | 0.77 | 0.62 | 0.63 | NA |
| GPP | 0.64 | 0.63 | 0.72 | 0.36 | 0.64 |
| NPP | 0.50 | 0.52 | 0.52 | 0.33 | 0.56 |
| ANPP | 0.45 | 0.46 | 0.42 | 0.28 | 0.48 |
| ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.64 |
| ANPP~woodystem~ | 0.17 | 0.23 | 0.12 | 0.19 | 0.28 |
| BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.35 |
| BNPP~fineroot~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
| R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.28 |

@@

With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes is much lower. Climate explains at most 23% of variation in ANPP~woody stem~, 16% in BNPP~fine root~, and 30% in R~root~. In contrast, climate strongly explains variation in ANPP~foliage~, with mean annual temperature explaining 63% of variation.

@@

Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses. We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux. However, field techniques for estimating components of primary productivity are variable in their accuracy. This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP. Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux. It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality. While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine root~, and, to some extent, NPP. We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change. It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.

>>>>

| Variable | Definition | Components included | Methodologies used | Number of records | Number of plots |
|--------------|-------------------|-------------|------------------------|-------|-------|
| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 257 | 80 |
| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 93 | 78 |
| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 258 | 166 |
| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 94 |
| ANPP~woody~ ~stem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 266 | 252 |
| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 107 | 87 |
| BNPP~fine~ ~root~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 90 | 77 |
| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 65 | 43 |

@@

*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes. To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component. The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux. Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015]. Stacked plots were generated for:  
1. GPP = NPP + R~auto~  
2. NPP = ANPP + BNPP  
3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
4. Total belowground carbon flux = BNPP + R~root~

*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~. For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated. The ratios were regressed against latitude and climate variables, using the linear model specified above. Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions. To account for this, models were re-run using only data from sites $\leq$ 1000m.

@@

[*Relationships and differences among fluxes- *] In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig. 2). That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.

@@

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.4). There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.

@@

We found a significant relationship between productivity and potential evapotranspiration for all fluxes. ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig. 5). We find strong evidence for a saturation point or peak with PET: productivity tends to increase at values below 1000mm, before saturating between 1200 and 1700mm. There is evidence that productivity begins to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of productivity for all fluxes. BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig. 5). Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa. At values above 0.8 kPa, productivity began to decrease.

@@

```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("combined_plots.png")
```

*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity. We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~. Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases (fig. 5). ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decrease linearly with temperature seasonality. Temperature seasonality is strongly correlated with annual temperature range, and, as expected, all fluxes show almost identical responses to it. Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

@@

*Within the growing season, how does productivity vary with climate?* Within growing season months, we find that climate has a much weaker effect on productivity. For each of temperature, precipitation, PET, and solar radiation, we find a small effect of climate for certain carbon fluxes. There is a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP~foliage~ (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05). Productivity increases with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine~ ~root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine~ ~root~ (R^2^ = 0.11, p<0.05), and ANPP~woody~ ~stem~ (R^2^ = 0.06, p<0.05).

@@

| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | MAT*MAP |
|----------|----------|----------|----------|--------------|----------|
| R~auto~ | 0.65 | 0.77 | 0.62 | 0.63 | NA |
| GPP | 0.64 | 0.63 | 0.72 | 0.36 | 0.64 |
| NPP | 0.50 | 0.52 | 0.52 | 0.33 | 0.56 |
| ANPP | 0.45 | 0.46 | 0.42 | 0.28 | 0.48 |
| ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.64 |
| ANPP~woody~ ~stem~ | 0.17 | 0.23 | 0.12 | 0.19 | 0.28 |
| BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.35 |
| BNPP~fine~ ~root~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
| R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.28 |

@@

With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes is much lower. Climate explains at most 23% of variation in ANPP~woody~ ~stem~, 16% in BNPP~fine~ ~root~, and 30% in R~root~. In contrast, climate strongly explains variation in ANPP~foliage~, with mean annual temperature explaining 63% of variation.

@@

Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses. We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux. However, field techniques for estimating components of primary productivity are variable in their accuracy. This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP. Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux. It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality. While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody~ ~stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine~ ~root~, and, to some extent, NPP. We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change. It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.
